### 2017-09-22 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Continued text, on reservoir part

		\subsection{Implementing Bayesian analysis numerically with Python and PyMC}
		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}. For doings this, the programming language of choice in this work is Python. The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}. Development is facilitated by an expressive but concise and clean syntax that is easy to learn. Python is dynamic, compatible with multiple platforms and offers good support for numerical computing. Integration of other scientific libraries and extension via C, C++, Fortran or Cython is easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}. Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.

		In probabilistic programming, programming variables are used as components to build probabilistic models \citep{davidson2015}.

		The numerical implementation of Bayesian analysis is further aided by the use of the Python library PyMC3, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \citep{davidson2015, salvatier2016pymc3}. Different model fitting techniques are provided in PyMC3, such as the \textit{maximum a posteriori} (MAP) method and several \textit{Monte Carlo Markov Chain} (MCMC) sampling methods \citep{salvatier2016pymc3}. 
		Name examples for state-of-the-art samplers included.
		\citet{salvatier2016pymc3} point out that the development of PyMC3 continuing, as the inclusion of further tools is planned for future updates.

		\subsection{Reservoir Volumetric Calculation}

>>>>

        \subsection{Economic significance and reservoir valuation}\label{sec:Reservoir_values}
        In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process. Changes in Bayes actions are considered as a measure for this effect.\\
        For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization. These are typically comprised stratigraphically of at least one potential reservoir formation, succeeded by at least one sealing layer above. Additionally, a trapping mechanism is normally required (source?).\\
        The resulting reservoirs are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e. Bayesian decision theory is applicable. In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.\\
        Before pressure and production tests have been conducted (i.e. before production has started), volumetric estimation is the only approach to assess the amount of hydrocarbons in place in a reservoir. From this value, recoverable reserves can be estimated based on an estimated recovery factor \citep{dean2007volumetric}.\\
        Oil-in-place and gas-in-place volumes are calculated based on:
        \begin{enumerate}
        \item Subsurface rock volume containing hydrocarbons. This is mainly defined by thickness and areal extend of the accumulation.
        \item Weighted average effective porosity of the reservoir rock.
        \item Water saturation in the reservoir rock.
        \item Hydrocarbon fluid properties \citep{dean2007volumetric}.
        \end{enumerate}
        As the cases used in this work are purely artificial, a pure oil accumulation is assumed. Using the factors listed above, the respective equation for original oil-in-place (OOIP) is formulated as follows:
        \begin{equation}\label{eq:OOIP}
        OOIP = A * h * \phi * (1 - S_W) * 1/FvF
        \end{equation}
        Where OOIP is returned in $m³$. The hydrocarbon-filled rock volume is defined by the drainage area $A$ in $m²$ and the net pay thickness $h$ in $m$. Porosity $\phi$ and water saturation $S_W$  (interstitial water) are given in fraction of the rock volume. A dimensionless factor for the change in oil volume between reservoir conditions and standard conditions at surface is represented by the formation volume factor $FvF$. Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric}.\\
        Subsequently, the effectively recoverable oil reserves can be calculated by multiplying the OOIP with a recovery factor $RF$. It is influenced by a number of fluid properties, such as viscosity, density, solution oil/gas ratio and the formation volume factor. Thus, it is difficult to estimate \citep{dean2007volumetric}. According to

        - in the end: NPV

        - here: 1D abstract case: arbitrarily chosen "scoring-system" as a reservoir value measure

		\subsection{Numerical computational implementation via Python, GemPy and PyMC}
		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}. The implicit method of forward geological modeling, described above, is to be embedded in such a framework. For doing so, the programming language of choice in this work is Python. The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}. Development is facilitated by an expressive but concise and clean syntax that is easy to learn. Python is dynamic, compatible with multiple platforms and offers good support for numerical computing. Integration of other scientific libraries and extension via C, C++, Fortran or Cython are easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}. Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.\\
		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citet{delaVarga2017gempy}. Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC (see below). At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
		For conducting the Bayesian analysis, GemPy is combined with the Python library PyMC, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \citep{davidson2015, salvatier2016pymc3}. Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc} \citep{salvatier2016pymc3}. The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC. the values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}. \textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}. Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} variables \citep{delaVarga2016}. Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\		
		As of writing this, the newer PyMC3 is not yet supported by GemPy. Thus, it is resorted to PyMC 2 for 3D geological settings in this work. However, the newer PyMC 3 is used in a simple 1D scenario. \citet{salvatier2016pymc3} point out that the development of PyMC 3 is continuing, as the inclusion of further tools is planned for future updates.
